---
title: "MGSC 310 Final Project"
author: "Luis Rivas, Joseph Calise, Jasper Wu, BJ Bae"
subtitle: MGSC 310 Final Project
output:
  html_document:
    df_print: paged
  html_notebook: default
---

```{r setup, include=FALSE}


library(knitr)

# As long as you are working in a Rstudio Project file, you shouldn't need to 'hard code' directories like this 
# change to your own working directory
#knitr::opts_knit$set(root.dir = 'C:/Users/doosti/Desktop/MGSC_310')
#knitr::opts_knit$set(root.dir = '/Users/luisrivas/Documents/CPSC_Courses/MGSC_310/')
#knitr::opts_knit$set(root.dir = '/Users/jasperwu/Chapman/Spring2023/MGSC310/MGSC_310/problem_sets')

# set seed to your own favorite number
set.seed(310)
options(width=70)
# if you want to prevent scientific format for numbers use this:
options(scipen=99)

# general rchunk code options
opts_chunk$set(tidy.opts=list(width.wrap=50),tidy=FALSE, size = "vsmall")
opts_chunk$set(message = FALSE,
               warning = FALSE,
               cache = TRUE,
               autodep = TRUE,
               cache.comments = FALSE,
               collapse = TRUE,
               fig.width = 5,  
               fig.height = 4,
               fig.align='center')

```

```{r setup_2}

# load all your libraries here
library('tidyverse')
library('dplyr')
library('rsample')
library('ggridges')
library('ggplot2')
library('tree')
library('caret')
library('randomForest')
library('caret')
library('randomForestExplainer')
library('ggcorrplot')
# note, do not run install.packages() inside a code chunk. install them in the console outside of a code chunk. 

```


## Exploring our dataset. 

```{r}

# code for data exploration 
apartments <- read_csv("../datasets/Equity_Apartments_Data.csv")
dim(apartments)
apartments <- apartments %>% select(Price, Beds, Baths, City, sq.ft, Floor, Units, Balcony, Walk_In_Closet, Fireplace, City_Skyline, Kitchen_Island, Stainless_Appliances, Office_Space, Renovated) 
apartments <- na.omit(apartments)
dim(apartments)

print(glimpse(apartments))

```
## Correlation plots

Evaluate the correlations among the continuous variables
```{r}

# correlation matrix
cor.table <- cor(apartments %>% select_if(is.numeric))
print(cor.table)
ggcorrplot(cor.table)

```
## Descriptive Summary of data - Price by City for apartments with 1 bed room and with >=2 beds

```{r}

table(apartments$City)

city_avg_bed2ormore <-
  apartments %>% filter(Beds>=2) %>% 
  group_by(City) %>%
  summarize(
    price_avg_city = mean (Price), na.rm = TRUE,
    num_apartment = n())
city_avg_bed2ormore %>% arrange(-price_avg_city)

city_avg_bed1 <-
  apartments %>% filter(Beds==1) %>% 
  group_by(City) %>%
  summarize(
    price_avg_city = mean (Price), na.rm = TRUE,
    num_apartment = n())
city_avg_bed1 %>% arrange(-price_avg_city)

```
## An interesting graph. 

It will be interesting to see how many different variables can be used to predict the rent price of apartments. 

```{r}

# Visualizing prices
print(ggplot(apartments, aes(x = Price)) + geom_histogram(bins = 50) + labs(title="Histogram of Rent Prices in America", x = "Rent Price", y = "Frequency") + theme_minimal()) 

```


## Model 1: Linear Regression

```{r}

# Model 1: Linear Regression
set.seed(310)
apartments_split <- initial_split(apartments, prop=0.8)
apartments_train <- training(apartments_split)
apartments_test <- testing(apartments_split)

mod1 <- lm(Price ~ Beds + Baths + sq.ft + Floor + City + Units + Balcony + Walk_In_Closet + Fireplace + City_Skyline + Kitchen_Island + Stainless_Appliances + Renovated + Office_Space, data=apartments_train)

summary(mod1)

# Model 1 Performance
preds_train <- predict(mod1, newdata = apartments_train)
preds_test <- predict(mod1, newdata = apartments_test)

RMSE(preds_train, apartments_train$Price)
RMSE(preds_test, apartments_test$Price)

# Model 1: Plotting Predicted Price vs Actual
results_test <- apartments_test %>% mutate(Price_pred = preds_test)

ggplot(results_test, aes(x = Price, y = Price_pred)) + geom_point() + 
  labs(x = "Actual Price", y = "Predicted Price", title = "Predicted vs Actual Price")

```


## Model 2: Random Forest

```{r}

# Model 2: Random Forest
mod2 <- randomForest(Price ~ Beds + Baths + sq.ft + Floor + City + Units + Balcony + Walk_In_Closet + Fireplace + City_Skyline + Kitchen_Island + Stainless_Appliances + Renovated + Office_Space,
                           data = apartments_train, 
                           ntree = 200,
                           mtry = 8, 
                           importance = TRUE)

summary(mod2)

# Model 2 Performance
preds_train <- predict(mod2, newdata = apartments_train)
preds_test <- predict(mod2, newdata = apartments_test)

RMSE(preds_train, apartments_train$Price)
RMSE(preds_test, apartments_test$Price)

# Model 2: Plotting Predicted Price vs Actual
results_test <- apartments_test %>% mutate(Price_pred = preds_test)

ggplot(results_test, aes(x = Price, y = Price_pred)) + geom_point() + 
  labs(x = "Actual Price", y = "Predicted Price", title = "Predicted vs Actual Price")

```


## Model Evaluation

### Model 2 Eval

```{r}

plot(mod2)

```

```{r}

importance(mod2)
varImpPlot(mod2)

```

```{r}

plot_min_depth_distribution(mod2)

```

```{r}

plot_predict_interaction(mod2)

```

```{r}

plot_multi_way_importance(mod2)

```